---
title: "Plot rms summaries and predictions"
author: "Richard Meitern"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot rms summaries and predictions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warnings = FALSE
)
library(rms)
```



### Load the reqired libraries
```{r setup}
#the regression modelling by F. Harrell
library(rms)

#the extension for plotting the rms summary objects
library(ormPlot)

#to modify plots
library(ggplot2)
```

### Looking at the bundeled data
OrmPlot has `educ_data` bundeled. `educ_data` is used in this vignette
to show the functionality of the package. To familiarize with the data:
```{r eval=FALSE}
#show first 6 rows
head(educ_data)
```


```{r echo=FALSE, results="asis"}
pander::pandoc.table(head(educ_data), split.tables=Inf)

```
```{r, eval = FALSE}
#show variable explanation
help(educ_data)
```

### Setting up rms datadist
The rms package requres that a `datadist` object is set up properly. According
to `datadist` documentation:
`q.effect` is a set of two quantiles for computing the range of continuous
variables to use in estimating regression effects.
Defaults are `c(.25,.75)`, which yields inter-quartile-range odds ratios
```{r}
#q.effect determines for what range the odds ratios are given on plots
dd <- datadist(educ_data, q.effect = c(0.5, 0.75))
#set it also to options
options(datadist="dd")
```

### Creating a model

```{r}
#see help(orm) for further info
orm_model<-orm(educ_3 ~ Rural + sex + n_siblings + cran_rzs + height_rzs + 
    FW_rzs + YOBc + (YOBc * sex) + (YOBc * Rural), data = educ_data)
```
### Plotting the model predictions

```{r}
plot(orm_model, "cran_rzs", "Rural")
```




### Plotting the model summary
The easiest way is to just plot the summary object
```{r fig.width=7, fig.height=4 }
plot(summary(orm_model))
```

If this does not look nice enogh you can also get `ggplot2` objects to customise to your needs.
The best way to get customizable plots is to specify `return_ggplots=TRUE`
```{r fig.width=6}
plots<-plot(summary(orm_model), return_ggplots=T)
plots[[1]]
plots[[2]]
```


These can be joined using the join_ggplots command. You can edit the plots as 
any ggplot plot
```{r, fig.width=7}

p1 <- plots[[1]] + 
  scale_x_discrete(labels=c("Mean", "Lower CI", "Upper CI"), 
                   position = "top",
                   name = NULL)

# the x axis is actually y axis because the cordinates are flipped with coord_flip()
p2 <- plots[[2]] + scale_y_continuous(breaks = c(0.5, 0.7, 0.9, 1.1),
                                position = "right")
  

forestplot<-join_ggplots(p1,p2)

```

To save as svg fur further editing just use `ggsave` from `ggplot2`
```{r, eval=FALSE}
ggsave("forestplot.svg",forestplot)
```

